---
title: "COVID Vaccine Signups: Sites Within 40 Minutes of Perkasie"
output:
  rmarkdown::html_document:
    theme: flatly
output_dir: "docs"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(jsonlite)
library(tidyr)
library(kableExtra)
library(data.table)
library(lubridate)
library(sf)
library(geojsonio)
library(remotes)
library(esri2sf)
library(mapboxapi)
library(tmap)

tmap_mode("view")

options(scipen = 999) 

options(timeout= 4000000) 
```

Notice: This is a personal website for research purposes. It is not intended to provide accurate information. Please use other resources for finding accurate information on COVID and do not share this link.

```{r, echo=FALSE, results = 'asis'}

today <- today()
cat("Data last updated ", format(today, format= "%B %d %Y"))
```

```{r, include = FALSE}

 # drive_40min <- mb_isochrone("Perkasie, PA",
 #                           profile = "driving",
 #                           time = 40)
 # 
 # saveRDS(drive_40min, "drive_40min_Perkasie.rds")

```

```{r, results = 'hide'}

drive_40min <- readRDS("drive_40min_Perkasie.rds") %>% 
  st_transform(3362)

url <- "https://services1.arcgis.com/Nifc7wlHaBPig3Q3/arcgis/rest/services/Vaccine_Provider_Information/FeatureServer/0"

vaccine_points <- esri2sf(url) %>% 
  st_transform(3362)

vaccine_points_in_drive <- st_filter(vaccine_points, drive_40min, .predicate = st_intersects) %>% 
  select(Facility_Name, Address, Phone_Number, Website)

unique_websites <- vaccine_points_in_drive %>% 
  st_drop_geometry() %>% 
  group_by(Website) %>% 
  summarise(Facility_Name = first(Facility_Name))

unique_phone <- vaccine_points_in_drive %>% 
  st_drop_geometry() %>% 
  group_by(Phone_Number) %>% 
  summarise(Facility_Name = first(Facility_Name))

```

## Unique Signup Websites Within 40 Minutes of Perkasie

This is a consolidated list of unique website signup links. Each link will include at least one signup facility within a 40 minute drive of Perkasie.

```{r}

unique_websites %>% 
  mutate(Website = cell_spec(Website, "html", link = Website, new_tab = TRUE)) %>% 
  kbl("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Manually Added Signup Websites

```{r}

names <- c("Wegmans Signup Link",
"Giant Pharmacy Signup (Local Giants)",
"The Other Kind of Giant Pharmacy Signup (further drive, still worthwhile)")

links <- c("https://www.wegmans.com/covid-vaccine-registration/",
           "https://giantsched.rxtouch.com/rbssched/program/covid19/Patient/Advisory",
           "https://covidinfo.reportsonline.com/covidinfo/GiantFood.html?queueittoken=e_giantfoodcovid19~q_cffbce75-75e1-4643-be70-ee6cd5d90735~ts_1612269313~ce_true~rt_safetynet~h_6021c466cb22ba819262fcba678c3655927c7a23e03a5e7ce338ab0667274b05")

manual_websites <- data.frame(names, links)

manual_websites %>% 
  mutate(names = cell_spec(names, "html", link = links, new_tab = TRUE)) %>% 
  select(names) %>% 
  kbl("html", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


## Unique Phone Numbers Within 40 Minutes of Perkasie

```{r}

unique_phone %>% 
  kbl() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## All Vaccine Sites within 40 Minutes of Perkasie

### Map of All Vaccine Sites

```{r, fig.width = 10, fig.height = 5}

popup_columns <- names(vaccine_points_in_drive %>% st_drop_geometry())

tm_shape(drive_40min) + tm_borders(lwd = 2) +
  tm_shape(vaccine_points_in_drive) + tm_dots(palette = "viridis",
                                            size = 0.08,
                                            popup.vars = popup_columns) +
  tm_basemap("CartoDB.Positron")

```

### Table of All Vaccine Sites

```{r}
vaccine_points_in_drive %>%
  st_drop_geometry() %>% 
  mutate(Website = cell_spec(Website, "html", link = Website)) %>% 
  kbl("html", escape = FALSE) %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

